---
title: "Use NMF to discover"
author: "Xun Zhu"
bibliography:
  - 'tutorial.bib'
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Not an Introduction to knitr}
output: knitr:::html_vignette
---

This simple tutorial will guide you through a typical analysis workflow with
NMFEM pipeline.  At the end of this tutorial the reader should be able to
do unsupervised clustering and important gene calling using non-negative matrix
factorization, and important gene module calling using the spin-glass based algorithm
FEM. Throughout the tutorial we will be using the mouse embryonic lung epithelial cell dataset
[@treutlein2014reconstructing].

# Unsupervised Clustering and Important gene calling

Let's begin by loading the `NMFEM` package:

```{r message=F, warning=F}
library(NMFEM)
```

and loading the dataset attached in the package:

```{r message=F, warning=F}
data(fpkm)
```

the unsupervised clustering and important genes calling is nicely packed in a single function. If
you are running this on a multi-core computer, feel free to set a higher `n_threads_` to take
advantage of parallelism. Depending on your hardware this might take while so grab a cup of coffee
while waiting for the results. :)

```{r message=F, warning=F}
nmf_results <- nmf_subpopulation(fpkm, n_threads_ = 30)
```

The `nmf_results` contains various dataframes and ggplot objects for different aspect of the analysis.
Let's take a look at a few important ones. We begin by looking at the clustering result:

```{r message=F, warning=F}
predict(nmf_results$nmf_result)
```

Then we can ask "what are the genes that provided the evidence for the separation of these two clusters?"

```{r message=F, warning=F}
nmf_results$gene_info
```

One way to understand the connection between the called clusters and the called important genes is that
these two clusters express two distinct expression patterns. That is, some genes are expressed highly
only in one cluster while some other genes are expressed highly only in the other cluster.

Let's see what these two expression patterns are:

```{r message=F, warning=F}
nmf_results$coef_line_plot
```

We can also see the D-score distribution of these two patterns. This plot also gives you an
idea the relative abundance of genes for each pattern.

```{r message=F, warning=F}
nmf_results$d_score_frequency_plot
```

There are also other interesting plots that allow various insights into the clusters, important genes,
and expression patterns. We encourage the reader to explore by looking into its struture (it's just a list):

```{r message=F, warning=F}
str(nmf_results, max.level=1)
```

By the way, all the ggplot objects in the list have dataframe counter-parts for easy access of the underlying
data. For example, `coef_line_dat` has the data for `coef_line_plot`:

```{r message=F, warning=F}
nmf_results$coef_line_dat
```

# Important Module generation

Again, everything you need is nicely packed in the function `spinglass_procedure`:

```{r message=F, warning=F}
module_results <- spinglass_procedure(fpkm, phe, leading_genes, mppi, 'mouse', n_threads_ = 30)
```

We can view the module graphs by:

```{r message=F, warning=F}
module_results$graph_plot
```

A summary containing important information about these modules can be found using:

```{r message=F, warning=F}
module_results$final_tb
```

## References